Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-03-23 referee_process/first_submission/referee_response.tex

\documentclass[aps, 11pt, singlecolumn]{revtex4-1} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\usepackage{natbib}
\bibliographystyle{apsrev}
\usepackage{setspace}

\begin{document}
%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\noindent
Dear Editor,
$\,$\newline

\newenvironment{myquotation}{
\begin{quotation}
\itshape
}{ 
\end{quotation}
}

\begin{singlespace}
We thank the two referees for their careful reading of our letter
and their detailed and constructive critique.
In responding to these two reports, the science within
our paper has been significantly strengthened.
Addressing the questions raised by report A led us to undertake
our first 3D runs, and we report on the results of these in the
revised draft.  In responding to report B,
we realized that our previous definition of the
Nusselt number (Nu) was flawed, and have since defined
a slightly modified, more appropriate Nu.  This new definition better unifies
the results at low and high Mach number (Ma) and makes it easier to understand
the changes between the low and high Ma regimes. We have streamlined our discussion
of the control parameters and have shown how evolved properties (e.g., Ma) depend
on the control parameters of our experiment.  Our previous presentation of the
experiment was confusing, and following the report's comments we feel that we
have improved the presentation.

@@

\begin{myquotation}
...However, it is true that the
interesting change in transport results appear when the flow becomes
supersonic, but this could mean high superadiabaticity (epsilon) *or*
high Ra (for fixed Pr). The results should all really be cast as
`at high enough Ra or high enough superadiabaticity, a high Ma flow
results and this changes the transport characteristics'.
Ultimately, the scaling results exhibited are in terms of Ra, and this
makes perfect sense, although these results are only confined to a
small paragraph at the end of the results section without much
elaboration or explanation. I personally would like a much more causal
relationship explained between the shock heating and `spinners'
and the heat transport results.
\end{myquotation}
In revisiting our results, we found that the substantial differences between
low and high $\epsilon$ in heat transport arose from a poor definition of
Nu and from changes to the overall stratification.  We have determined a
more consistent version of Nu (based on evolved properties rather than initial
properties).  Now, high and low $\epsilon$ transport is substantially similar.
We do see a small change in transport when 2D simulations achieve a mean
Ma of O(1), and to test this we have conducted an additional suite of
experiments at higher $\epsilon = 1$.  In these solutions, all of the flow
properties show a change of behavior at the point at which the flows enter the
sonic regime.  This is shown in Figs. 1, 3, and 4 and discussed in the text.

Figure 4 shows a surprising result.  We have removed the former Fig. 3 (fluxes
as a function of height) so that we can discuss the result in our current
Fig. 4 fully.  We find that the density contrast across the domain at high $\epsilon$
appreciably shrinks in the evolved state compared to the initial state.  Furthermore, we find
that density inversions form near the boundary layers of these atmospheres once the mean atmospheric Mach number approaches
one, and show that this density inversion becomes significant at high Ra and high $\epsilon$,
and persists in 3D.  This is discussed fully in the Results \&
Discussion section.  We have combined the Results \& Discussion sections in the revised draft.

@@

Eqn (1): Better define co-ordinates, especially $z$ and $z_0$
\end{myquotation}
We agree that $z_0$ was confusing notation and this has been removed.  
We define $z$ shortly below Eqn (1).

\begin{myquotation}
The non-dimensionalization is very confusingly written. Can you write
out the non-dimensionalised polytope? This section also says that the
`timescale is the isothermal sound crossing time of the layer'
and then two sentences later says that `we use buoyancy time
units', so which is it? The definition of the latter does look like
the sqrt(epsilon) so maybe these are the same?
\end{myquotation}
We have clarified that the equations are non-dimensionalized on
the isothermal sound crossing timescale.  We have also clarified
how the evolution time of the flow is linked to the buoyancy time.
This appears in the final paragraph of the Experiment section.

\begin{myquotation}
`The scaling of the entropy gradient with epsilon ... evolved
values ': I really have no idea what these two sentences mean,
sorry!
\end{myquotation}
We agree that this sentence was confusing and have moved the
discussion of the magnitude of evolved fluctuations to the
discussion of Fig. 1.

\begin{myquotation}
Eqns: I don't understand why these equations are formulated with
gradients of nu when mu is constant. The stress tensor only depends on
mu and so this can be pulled out of all derivatives. Or am I missing
something? The case of constant nu and therefore variable mu is much
harder. Similarly for the formulation in terms of chi and not the
thermal conductivity; variable thermal conductivity is hard but
variable chi is easy. Why not write in terms of mu and k not nu and
chi?
\end{myquotation}
Our equations are formulated as the evolution of velocities rather
than momentum; this formluation is crucial to our current IMEX
timestepping of acoustic waves at low Mach numbers.  
In this velocity-based formulation, $\nu$ and $\chi$ are the
natural variables.  If $\nu$ and $\chi$ are constant in space,
the problem becomes a constant coefficient problem and is numerically
simple.  However, a spatially constant $\nu$ and $\chi$ does not satisfy
thermal equilibrium within the initial polytrope, which is only satisfied when
$\kappa_0 = \rho_0\chi$ is constant (where as in the letter, subscript 0
denotes initial values).  To satisfy thermal equilibrium requires
a gradient in $\chi$, and to satisfy a constant Pr through the domain we
must also have a gradient in $\nu$.  $\nu$ and $\chi$ are now nonconstant
coefficients, and this approach, as we take here, is much more challenging numerically. 
 We have
clarified and detailed the profiles of $\nu, \,\chi, \,\mu,\text{ and }\kappa$
in the second paragraph of the Experiment section.

@@

\begin{myquotation}
`IMEX': I am assuming this acronym stands for implicit-explicit?
Maybe write out?
\end{myquotation}
We have spelled this out in the second to last paragraph of the Experiment section.

\begin{myquotation}
`extended to include variable nu and chi' : see above. This seems
very unnecessary!
\end{myquotation}
See above description of $\nu$ and $\chi$ in our equations.

\begin{myquotation}
Results

@@

$d_z T_{ad} = 0$: This is only true for liquids not gases. See Speigel and
Veronis 1960.
\end{myquotation}
This is a good point that we had not appreciated.  Most of the studies of Boussinesq
convection with which we are familiar make the assumption that the fluid is incompressible.
It is true that in an ideal gas which is compressible but barely stratified, the
adiabatic temperature gradient is what we found and what is reported by
Spiegel and Veronis, $\partial_z T_{ad} = -g/c_P$ \cite{spiegel&veronis1960}.
We have updated the text following Eqn 8 to reflect this.

\begin{myquotation}
`While it has been suggested that pressure forces...': I do not
understand the discussion here. Do the authors regard the breakup of
the down flows as an extreme version of asymmetry?
\end{myquotation}
On further reflection, we realized that we are uncertain about how to
quantify asymmetries as the flows break up.  
Being unable to determine how $\epsilon$ affects symmetries in the
flow, we have removed this discussion.

@@

\begin{myquotation}
`exhibit states in which the flux entering...': Surely in a
stationary state the flux in has to equal the flux out. Are the
authors saying that this is not a stationary state?
\end{myquotation}
As detailed in the revised text, the combination of fixed-temperature
boundary conditions and no-slip boundary conditions allows the system to alternate between
roll states of convection and shear states of convection.  These states persist over long
time scales (hundreds of buoyancy times) and the atmosphere is in flux disequilibrium in both states.
A sensible equilibrium and average emerges over long times.  This discussion appears with
the discussion of Fig. 3a.

\begin{myquotation}
Discussion

Can the authors demonstrate the basic principles of whatever balance
is described by the Grossman-Lohse theory?
\end{myquotation}
We believe that the Grossman-Lohse theory, which models the energy dissipation rate of both
kinetic and potential energies, could be extended to stratified, compressible flows which also
have potential energy.  At low $\epsilon$, it seems the extension would be easier to
develop. However, we are inexperienced in boundary layer theory and presently 
cannot say with certainty that an expansion of this theory to stratified convection is tractable. 
We have removed this suggestion from letter.

### 2017-06-12 referee_process/second_submission/referee_response.tex

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[aps, 11pt, singlecolumn]{revtex4-1} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\usepackage{natbib}
\bibliographystyle{apsrev}
\usepackage{setspace}

\begin{document}
%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\noindent
Dear Editor,
$\,$\newline

\newenvironment{myquotation}{
\begin{quotation}
\itshape
}{ 
\end{quotation}
}

\begin{singlespace}
Editor letter goes here

$\,$\newline
\noindent
Sincerely,

Evan H. Anders \& Benjamin P. Brown

%\end{singlespace}

$\,$
\newline
$\,$
\newline
\noindent
\Large{\textbf{Response to report B:}}\newline$\,$\newline\indent

\begin{myquotation}

The second submission of this paper, after the first round of review,
is SUBSTANTIALLY different from the first time around, and essentially
needs refereeing from scratch as a new paper. The paper does still
essentially examine the heat transport in compressible convection as
as function of superadiabaticity and Rayleigh number (and therefore
varying thermal diffusivity) for fixed density stratification and
Prandtl number via 2D simulations The new paper adjusted the
formulation of the Nusselt number, changing the results substantially,
and added some 3D simulations to the previously all-2D simulations.
The major result is the scalings of the heat transport (measured by
their Nusselt number) with the Rayleigh number bears a remarkable
similarity to standard Boussinesq Rayleigh-Benard convection, at all
the wide range of superadiabaticities simulated. This in itself is an
interesting result, as it somewhat surprising considering that their
fixed background stratification has quite a strong density contrast
($\sim$20) and some of their departures from adiabaticity are large
($\epsilon \sim$ 1 $\rightarrow$ adiabatic index m=0.5).

I think this paper is getting closer to publishable, although I still
find the explanation of the setup of the model in particular somewhat
opaque though, and the interesting results not really addressed in any
detail, as I'll explain below.

I am glad that the authors took my comments on the setup from before
at least partially to heart! I still think things could be clearer.
Most specifically again, is the question of what are the salient
parameters of compressible convection. There are 4, as mentioned
previously, and the authors here have chosen to keep one fixed -- a
measure of the stratification, for which they use $n_\rho$. For the
others, it is nice to cast things in terms of a Rayleigh number and a
Prandtl number since this gels nicely with standard Boussinesq
Rayleigh-Benard convection (RBC). However, then the third parameter,
the superadiabaticity, which the authors call $\epsilon$, is also part of
the definition of the Rayleigh number, and therefore, as it should be,
the Rayleigh parameter is really a definition of the thermal
diffusivity. To the unaware reader more familiar with RBC, this can be
a bit confusing, since there are no independent measures of the
driving (superadiabaticity) and the thermal diffusivity there. It
would really help the reader here to point out the following in
relation to the simulation sets that are performed:

\begin{itemize}
\item At fixed $\epsilon$, varying Ra means that the thermal diffusivity $\chi_t$
scales like 1/$\sqrt{\text{Ra}}$.
\item At fixed Ra, varying $\epsilon$ means that the thermal diffusivity $\chi_t$
scales like $\epsilon$.
\item Since Pr=fixed, viscous diffusivity scales like thermal diffusivity.
\item (and of course, all vary with depth individually)
\end{itemize}
\end{myquotation}
I need to add this to the experiment section.

\begin{myquotation}
Beyond that, I think everything is right. The expression for the
polytopes in terms of the number of density scale heights makes the
notation a little over-complex. There are some disconnects in where
the non-dimensionalisation is performed (they say that ``take R=1'' at
one point and then non-dimensionalise R out again later, for example).
The description of which of the diffusivities or the
conductivity/dynamic viscosity are constant could certainly be
tightened up.
\end{myquotation}
I need to tighten up the description of polytropes, and only talk about
non-dimensionalizing R once.

\begin{myquotation}
With regard to the results:

In light of what I said above, the results in Fig 2. are not
surprising. Higher Ra here at fixed epsilon means lower thermal
diffusivity, and therefore eddies might be expected to retain their
identity longer. The above explanation would help.
\end{myquotation}
Agreed, we hope that the added explanation makes this clearer to the
reader.

\begin{myquotation}
There is quite a bit of mention of “windy” states without any real
technical description of what they are. Please either give more
information (at least something visual to distinguish from non-windy
states) or remove the distraction.
\end{myquotation}
I think it might be a good idea to remove them in order to keep it 
letter-length, but I'll see if I can keep it short \& explain 
relevant pieces.

\begin{myquotation}
Nusselt number: Usually this is a ratio of heat transport in the
turbulent state to that in the conduction state. So therefore, aren’t
the two $F_A$’s different on the top and the bottom? The top has a
modified kappa but the bottom has the original kappa profile?
\end{myquotation}
Nope!  That's our new change.  Maybe make this clearer.

\begin{myquotation}
I think the paragraphs on the Nu vs Ra and the Re/Pe vs Ra are the
meat of the paper! It would really be nice here to know what causes
the difference between 2D and 3D in the Nu plots. The 2/7 law is often
associated with more “windy” states. That seems LESS likely in 3D, so
what is going on? Note also that the sensitive dependence on the exact
roll or other structure seems to imply that the simulation box is too
small. This dependence should not be the case.
\end{myquotation}
Yeah, AR=4, 8, and 16 all return the windy states, so there's some strange
physics here.  I'll try to explain the discrepancy between 2D and 3D, though.

\begin{myquotation}
The Re vs Ra are a bit mysterious at low Ra. I would expect the Re ~
RA$^{1/2}$. Can you explain these alternative scalings or at least why the
expected scaling emerges at large Ra? Are the 2D ones just
over-constrained?
\end{myquotation}
Yes, can explain: we're combining the 1/2 scaling law that we expect from
the Ra changing diffusivity with the 1/4 scaling law of rising Ma in 2D,
which produces the 3/4 law.

\begin{myquotation}
Regarding Fig 4 — are the authors calculating the integrated evolved
density profile to get this number? If the effect is only in the
boundary layer as they mention, I am surprised that the deviation is
so large. This figure needs much more explanation!
\end{myquotation}
We're calculating the horizontally averaged, time-integrated, evolved density
profile and then taking two measures of it.  One is max/min, one is top of domain
/ bottom of domain.  We should probably explain this more and mention that it persists
even at high $\epsilon$ in 3D.

\noindent
This concludes our response to report B.  We thank referee B for this second report.

$\,$
\newline
$\,$
\newline
\noindent
\Large{\textbf{Response to report C:}}\newline$\,$\newline\indent

\begin{myquotation}
This Letter reports results from a numerical study of thermal
convection in a compressible fluid at unity Prandtl number and a range
of initial stratification. The main result is the scaling of the
Nusselt number with the Rayleigh number, supplemented by some
observations such as characteristic changes in the transition between
subsonic and supersonic regimes.

This problem is highly challenging. Ideally, a numerical exploration
should be guided by some theoretical insights or practical
observations. The present paper appears to be ``purely'' numerical
without such guidance, albeit some comparison with the well-known
problem of incompressible Rayleigh--Bernard convection. Nevertheless,
the results are significant and should be published. While I do not
wholeheartedly recommend this paper to PRL, I would not object to its
acceptance.
\end{myquotation}
Is there any way we can bring in some more ``we're guided by X and thus
we expected to get our results''?

\begin{myquotation}
Below is a minor comment that the authors may find useful.

Most parts of the paper is relatively well written, except in the
introduction. Here, there is room for improvement. The introductory
paragraph is virtually empty. I would cite a couple of papers for
"These prior studies" and be specific about both "important insight"
and known "fundamental properties". The final paragraph of the
introduction has failed to fulfill its mission. I would explicitly
state (at least) the most significant result of the paper in this
paragraph.
\end{myquotation}
We need to beef up the intro, as specified here.

$\,$\newline\noindent
This concludes our response to report C.  We thank referee C for this report.

$\,$
\newline
\noindent

\end{singlespace}

\bibliography{../../biblio.bib}
\end{document}

### 2017-06-12 tex/nusselt_paper.tex
reworks the Ra/Pr discussion

Convective dynamics are controlled by the superadiabaticity
of the atmospheres as well as the atmospheric diffusivities.
The superadiabaticity, or the magnitude of the (negative) 
entropy gradient, is set by 
the superadiabatic excess, $\epsilon = m_{ad} - m$, where 
$m_{ad} = (\gamma - 1)^{-1}$.
The atmospheric thermal diffusivity, $\chi$,
and kinematic viscosity (viscous diffusivity), $\nu$,
are determined by the Rayleigh number (Ra) and the Prandtl
number (Pr).  We set $\text{Pr } = \nu/\chi = 1$ throughout
this work. The polytropic initial conditions are in
thermal equilibrium, $\kappa_0 \partial_z T_0 =$ const,
so $\kappa_0 = \chi \rho_0$ must be constant. To keep
Pr constant with height, we set $\chi = \chi_t/\rho_0$
and $\nu = \nu_t/\rho_0$.  Under these constraints, the
diffusivity profiles are specified by choosing Ra
at the top of the domain,
\begin{equation}
\text{Ra}_{\text{top}} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\end{equation}
where $\Delta S_0 = \epsilon\ln (1 + L_z)$ is the initial entropy difference 
between the top and bottom of the atmosphere and
$c_P = \gamma(\gamma-1)^{-1}$ is the specific heat 
at constant pressure. The profiles of $\nu$ and $\chi$ are constant with
time, and Ra at the bottom of the atmosphere is greater than Ra$_{\text{top}}$
by a factor of $e^{2n_{\rho}}$.  This formulation leaves
the thermal conductivity, $\kappa = \rho\chi$, and
the dynamic viscosity, $\mu = \rho\nu$, free to evolve
as the density profile changes.  Henceforth,
when we use Ra, we are referring to Ra$_{\text{top}}$.

@@

We impose impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $T_0$ and $\epsilon$.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power. We nondimensionalize our computational domains by setting
all thermodynamic variables to unity at $z = L_z$, choosing
$R = T_t = \rho_t = 1$.  By this choice, the non-dimensional
length scale is the inverse temperature gradient scale and the 
timescale is the isothermal sound crossing time, 
$\tau_I$, of this unit length.
Meaningful convective dynamics occur on 
timescales of the atmospheric buoyancy time,
$t_b = \tau_I \sqrt{L_z/g\epsilon}$. 
All reported results are taken from time averages
over many $t_b$ beginning \{100, 40\}$t_b$
after the start of our \{2D, 3D\} simulations in order to
assure our results are not biased by the convective transient.

>>>>

Convective dynamics are controlled by the superadiabaticity
of the atmospheres as well as the atmospheric diffusivities.
The superadiabaticity, or the magnitude of the (negative) 
entropy gradient, is set by 
the superadiabatic excess, $\epsilon = m_{ad} - m$, where 
$m_{ad} = (\gamma - 1)^{-1}$ and the initial entropy
difference between the top and bottom boundaries is
$\Delta S_0 = \epsilon\ln (1 + L_z)$.  At a fixed value of
$\epsilon$, the diffusivities are set by the
Rayleigh number (Ra) and the Prandtl number (Pr),
\begin{equation}
\text{Ra}_{t} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\qquad
\text{Pr} = \frac{\nu}{\chi},
\end{equation}
where $c_P = \gamma/(\gamma-1)$ is the specific heat at
constant pressure, and $\nu$ and $\chi$ are the thermal diffusivity and
kinematic viscosity, respectively.
The atmospheric thermal diffusivity, $\chi$,
and kinematic viscosity (viscous diffusivity), $\nu$,
are determined by the Rayleigh number (Ra) and the Prandtl
number (Pr).  These polytropes are initially in
thermal equilibrium, so the initial thermal
conductivity, $\kappa_0 = \chi \rho_0$ must be
constant.  Throughout this work we specify
that Pr $= 1$ and is constant with depth.  By this
choice, $\chi = \chi_t / \rho_0$ and $\nu = \nu_t / \rho_0$.
We leave $\kappa = \rho\chi$ and the dynamic viscosity,
$\mu = \rho\nu$, free to evolve as the density 
profile evolves.  This formulation 
means Ra at the bottom of the domain is greater than
Ra$_t$ by a factor of $e^{2n_\rho}$, and henceforth
when we specify Ra we are referring to Ra$_t$.

Our diffusivities thus scale as
\begin{equation}
\chi_t \propto \sqrt{\frac{\epsilon}{\text{Pr}\cdot\text{Ra}_t}}, \qquad
\nu_t \propto \sqrt{\frac{\epsilon\cdot\text{Pr}}{\text{Ra}_t}},
\end{equation}
and for constant Pr = 1, as in this work, $\chi_t = \nu_t$.

@@

We impose impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain such that 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $T_0$ and $\epsilon$.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power. We nondimensionalize our computational domains by setting
all thermodynamic variables to unity at $z = L_z$, choosing
$T_t = \rho_t = 1$.  By this choice, the non-dimensional
length scale is the inverse temperature gradient scale and the 
timescale is the isothermal sound crossing time, 
$\tau_I$, of this unit length.
Meaningful convective dynamics occur on 
timescales of the atmospheric buoyancy time,
$t_b = \tau_I \sqrt{L_z/g\epsilon}$. 
All reported results are taken from time averages
over many $t_b$ beginning \{100, 40\}$t_b$
after the start of our \{2D, 3D\} simulations in order to
assure our results are not biased by the convective transient.
